Collaborative Learning with Dirichlet Process Clustering for Rapid Online Adaptation in Robotics

### 2023-09-05 paper/paper.tex
Sep 5

\usepackage{bm}
%\usepackage{amssymb}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\renewcommand{\algorithmiccomment}[1]{\hfill $\triangleright$ #1}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\hypersetup{hidelinks}
%\usepackage[colorlinks, linkcolor=blue, citecolor=green, urlcolor=red]{hyperref}

@@

%\begin{multicols}{2}

\section{Introduction}
Locomotion controls of legged robots are typically too complex to be designed manually.
In recent years, deep Reinforcement-Learning (deep RL) has made huge success in game playing
\cite{alphaGo, alphaStar} and is regarded promising for application in robotics. However, training such well performing agents requires enormous amounts of interactions with the environment, meaning that this can only 
be done in simulation where we can model the interaction and collect the training data much faster than in real-time. This raises a problem for application in robotics since there is always a gap between simulation and reality, and the policy trained in simulation could suffer from this sim-to-real transfer. Another problem is that any unforeseen changes to the environment dynamics could lead to severely degraded performance. These deep RL trained agents require extra fine-tuning or even retaining to retain their capabilities.
For example, the team of OpenAI Five \cite{openAI5} had to conduct three additional amendments in the architecture followed by extra trainings to incorporate small version updates in Dota 2 during the development of the agent.
For game-playing agents, these drawbacks are tolerable as there is no sim-to-real problem, and the environment is overall stationary. For robotics however, sim-to-real must be considered, and the environment dynamics is typically non-stationary as the robot can be travelling between different terrain, carrying different payloads or suffering from damages in its hardware.

It would be better if we may learn the true dynamics from the interactions with the current dynamics and
correct our policy accordingly. This is a model-based reinforcement learning (MBRL) solution that allows robots to learn policies with lesser interactions by learning a dynamics model and use it optimize the policy \cite{MBRL, black_box_search, policy_search}. 
However, the amount of data required to learn a model typically scales exponentially with the dimensionality of the input space \cite{curse_of_dim}, hence it is not suitable for online adaptation. 
The promising approach to address this is to use a repertoire-based method, where we learn a repertoire of elementary policies (e.g., one policy for turning left, one for moving forward, etc.) using quality diversity optimisation \cite{QD} to fully cover the task space (e.g., the 2D displacement we want the robot to make). 
The repertoire-based method is a hierarchical control method that keeps a variety of elementary policies and treats each policy as an action. 
We then give each elementary policy a behaviour descriptor (BD) to quantify the way it acts.
This enables us to skip the low-level states and actions (i.e., joint encoders and torques for all motors) and to work with the high-level behaviour space that has much smaller dimensionality.
During the adaptation, we learn the distortion (called transformation model) in the behaviour space from real-world interactions. Thanks to the reduced dimensionality, the transformation model can be learned in real-time.
Instead of optimising the policy with this learned model, we use a repertoire-based control that first predicts the outcomes for each elementary policy in the repertoire and then simply chooses the one with the predicted outcome that is most aligned with the current goal. 
This is made available for keeping a repertoire of policies rather than a single policy, so that the robot can adapt to the distortion by finding the corresponding policies that compensate the distortion (for example the robot aims to go forward and distortion is left, then a policy originally aiming right-front can be used for going forward instead).

The repertoire-based method has achieved very impressing results in task-solving \cite{EvoRBC} and online damage recovery \cite{RTE}. 
But this method heavily relies on having a good model that is both data efficient and suitable for modelling the distortion. 
The most commonly used model is Gaussian Process (GP) \cite{GP}. The prior mean function and the kernel are the most important factors of GP. 
If no prior knowledge is given about the dynamics, these  have to be selected manually according to experience.
If the distortion is very large (like broken legs) or complex, the GPs can be very inefficient or even misleading.
If we have data of real-world interactions across several different environments, we can build a GP for each of the situations and then teach the robot to identify the one that mostly explains its current situation. However, such real-world data are expensive to collect and could lead to covariate shift (the collected data distribution is different from that during deployment).
A promising solution that is to leverage the historical data of real-world interactions collected during deployment. 
After performing each task, the interaction data can be uploaded to an archive system where they will be analysed. 
Thus, the data doesn't have to be collected on purpose, and the acquired data distribution is completely unbiased. 
This also allows the robots to learn to adapt in a collaborative manner.
If we have multiple robots performing different tasks in different environments, we can quickly get an overview of the  common types of dynamics and build the adaptation strategy accordingly. 
If a robot encounters a new environment, with the interaction data uploaded, all the robots will be able to quickly adapt to this environment.
Finally, considering the scalability of such system, we can group the data into clusters based on their dynamics and assign a GP (with the corresponding mean and kernel) to each cluster. This prevents the adaptation strategy from becoming computationally expensive after we have collected data from thousands of tasks.

There are many challenges needs to be addressed to enable this collaborative learning. First, traditional clustering algorithms like k-nearest neighbours \cite{knn} and k-means \cite{k-means} need to specify the number of clusters. While in our case, this number is not even fixed as we might observe more clusters as we collect more data.
Second, we know that each task corresponds to a single environment, and hence we are essentially clustering the tasks instead of clustering the data points.
However, since some tasks are tougher than others, different tasks generate different amount of data. 
As a result, we will be clustering data of inconsistent sizes.
Third, since there is a large repertoire of policies, only a small subset of them will be executed while performing each task. Hence, it is very likely that the data collected from two tasks have no overlapping policies, making it hard to determine whether they have similar dynamics and whether they should be grouped together.
Moreover, after we have found the clusters, the data in each cluster can never cover the entire repertoire. Hence it is difficult to determine the prior mean for the policies that are not present in this cluster.
This paper aims to solve all the above problems by taking an non-parametric clustering method and using a uniquely designed prior mean function.

@@

\begin{algorithm}
\caption{MAP-Elites}
\begin{algorithmic}
\STATE \textbf{procedure} MAP-Elites
\STATE Discretize the task space into grids
\STATE $\mathcal{R} \leftarrow \emptyset$ \COMMENT{Initialize repertoire}
\FOR{$\bm{i} = 1 \rightarrow G$}
\STATE $\theta \leftarrow $ generate{\_}random{\_}policy()
\STATE $\Theta$, $f$ $\leftarrow$ evaluate{\_}policy($\theta$) \COMMENT{Get outcome and fitness}
\STATE add{\_}to{\_}repertoire($\Theta$, $f$, $\theta$, $\mathcal{R}$) \COMMENT{Update repertoire}
\ENDFOR
\FOR{$\bm{i} = 1 \rightarrow T$}
\STATE $\vartheta \leftarrow$ select{\_}parents($\mathcal{R}$)
\STATE $\bm{\theta} \leftarrow$ crossover($\vartheta$, $N$) \COMMENT{Generate $N$ off-springs}
\FOR{$\bm{j} = 1 \rightarrow N$}
\STATE $\Theta$, $f$ $\leftarrow$ evaluate{\_}policy($\bm{\theta_j}$)
\COMMENT{Off-spring evaluation}
\STATE add{\_}to{\_}repertoire($\Theta$, $f$, $\bm{\theta_j}$, $\mathcal{R}$)
\ENDFOR
\ENDFOR
\end{algorithmic}
\end{algorithm}
\noindent
Evaluating the policies in the simulation is the major source of computationally expense for the MAP-Elites.
A merit of this algorithm is that the inner for loop of off-spring evaluation can be conducted in parallel, and the repertoire update can be made after having collected all the results of the off-springs.
This enables us to generate a large number of off-springs in each iteration and evaluate them in parallel, fully leveraging the power of modern Cluster computing \cite{Q-Dax}.
By running the algorithm for a few thousand iterations, we will end up having a repertoire of high-performing policies well-covering the task space.

\subsection{Gaussian Process}
The Gaussian Process (GP) is a powerful non-parametric machine learning algorithm. 
It assumes the prior distribution of the outcomes to be a multivariate normal distribution with a prior mean $\mu$ and a covariance matrix $\Sigma$. 
\begin{equation}
f(\bm{y}) = \frac{exp(
-\frac{1}{2}(\bm{y} - \bm{\mu})^T \bm{\Sigma^{-1}} (\bm{y} - \bm{\mu})
)}
{
\sqrt{(2\pi)^N det(\bm{\Sigma})}
}
\label{multi_normal}
\end{equation}
In contrast to regular multivariate normal distribution, GP allows the prior mean and the covariance matrix to be input dependent. The
prior mean becomes a function $\mu(\bm{x})$, and the covariance between two data points is given by a covariance function 
$k(\bm{x_1}, \bm{x_2})$ called kernel. 
Following the fact that data points closer to each other are more correlated, this covariance starts from the prior variance and asymptotically decreases to zero as the distance between the two points tends to infinity. 
To account for the fact the observations might be noisy, we can incorporate the noise into the covariance matrix by adding the variance of the noise on the diagonal.
\begin{equation}
\bm{\Sigma} = \bm{K}(\bm{x_{1:N}}, \bm{x_{1:N}}) + \sigma_n^2 \mathbf{I}
\label{covariance_matrix}
\end{equation}

This allows us to build reasonable joint prior distribution for the values of any well-behaved functions.
To make prediction with GP, we first use the prior mean function and kernel to build up the joint distribution as the prior. 
Then we can calculate the posterior for the data we wish to evaluate conditioned on the ones we have observed using Bayes rule.
Since the Gaussian distribution is a conjugate distribution, the posterior is still a Gaussian with mean and variance given by \cite{GP, GP_posterior}:
\begin{equation}
\mu_*(\bm{x}) = \mu(\bm{x}) + 
\bm{\Sigma_{N, x}}^T
\bm{\Sigma_{N, N}}^{-1}
(\bm{y_{1:N}} - \bm{\mu}(\bm{x_{1:N}}))
\label{posterior_mean}
\end{equation}
%
\begin{equation}
\sigma^2_*(\bm{x}) = \sigma^2(\bm{x}) -  
\bm{\Sigma_{N, x}}^T
\bm{\Sigma_{N, N}}^{-1}
\bm{\Sigma_{N, x}}
\label{posterior_var}
\end{equation}
where $\bm{\Sigma_{N, N}}$ denotes the covariance matrix of the observed data, and $\bm{\Sigma_{N, x}}$ denotes the column vector with entry on each row equal to the covariance between 
each observed data and the data we hope to evaluate.

@@

\section{Methodology}
To formulate the problem clearly, we are aiming to leverage the collaboratively collected real-world interactions to assist a robot deployed in an unknown environment to quickly adapt and complete its task.
We are provided with some episodes (each task-solving process corresponds to an episode of interaction data) of real-world interactions collected by several robots deployed in an unknown distribution of environments.
The most intuitive idea is to label these historical data with their sources (like collected from grass terrain, collected with left leg broken, etc.) and build a simple database in the cloud server. 
During the adaptation, the robot examines the current environment and uploads the result to the database server. The server then identifies the most similar environment recorded in the database and let the robot download the corresponding data to assist the its adaptation.
However, accurately measuring and describing the environment can be expensive and difficult. It is also hard to determine whether a previously encountered environment shares similar dynamics with the current one even if we have accurate descriptions for them.
In the case we do know that some historical data come from exactly the same environment as the current one, these data might be insufficient in quantity to assist the current task.
Hence, it is reasonable to group the episodes resulting from similar dynamics into clusters, ensuring having enough data for analysis.
We will assume that we don’t have access to any information of the environment and aim to conduct clustering only according to the dynamics.

Previous works of clustering GP via DP have been made in \cite{infinite_MGP, variational_MGP}. 
But they work with individual data points, while in our case all the interaction data collected from the same task share the same indicator.
In this paper, we introduced a novel method to cluster batched data with infinite mixture of GPs while ensuring high flexibility.

There are a few assumptions needs to be made for further analysis. The first assumption is that the simulated results always correspond to real-world outcomes in some environment. 
For example, we set the friction coefficient to 0.9 in the simulation, but the it underestimated friction and the simulated outcomes are very close to the real-world results for friction coefficient equal to 0.8.
Hence we believe that although the simulated results are not accurate, they are still close to some real-world results.
This is not always true, but it plays a very important role in the methodology.

$$we \, have \, talked \, about \, prior \, mean \, importance$$
$$we \, have \, talked \, about \, simulation \, assumptions$$

\subsection{Linear Prior Mean Function}
The prior mean function gives a prior estimate of the dynamics for a certain environment.
This means we need to estimate the outcomes of every elementary policy in this environment.
The dynamics can efficiently expressed in the following form:
x_d(\bm{\beta}, \bm{\epsilon}) = f_d(\bm{\beta}, \bm{\epsilon}) 
+ g_d(\bm{\beta}) 
+ h_d(\bm{\epsilon}) + c_d 
\label{dynamics_assumption}
where the $x_d$ stands for the $d^{\text{th}}$ dimension of the policy rollout outcome (for example, this can be the robot displacement in the $x$ axis). Vector $\bm{\beta}$ represents the behaviour descriptor of the controller; vector $\bm{\epsilon}$ is the environment descriptor, representing some quantities of the environment. 
It can be seen from Eq. (\ref{dynamics_assumption}) that the dynamics consists of a term that depends on both the policy behaviour and the environment plus the terms that only depends on either the behaviour or environment plus a constant $c_d$.
Note that we do not require access to the $\bm{\beta}$ and $\bm{\epsilon}$ or knowing their detailed meanings.
Eq. (\ref{dynamics_assumption}) is generally non-linear. But in the case where the dynamics is relatively simple, like the use of periodic controller in RTE, we can get an approximation of it by Taylor expanding the $f_d(\bm{\beta}, \bm{\epsilon})$ term to the second order:
f_d(\bm{\beta}, \bm{\epsilon}) \approx
f_d(\bm{\beta}_0, \bm{\epsilon}_0) + 
\begin{bmatrix}
J_{\bm{0, \beta}} & J_{\bm{0, \epsilon}} \\
\end{bmatrix}
\begin{bmatrix}
\Delta \bm{\beta} \\
\Delta \bm{\epsilon}
\end{bmatrix}
+ \frac{1}{2}
\begin{bmatrix}
\Delta \bm{\beta}^T & \Delta \bm{\epsilon}^T \\
\end{bmatrix}
\begin{bmatrix}
H_{\bm{0, \beta \beta}} & H_{\bm{0}, \bm{\beta \epsilon}} \\
H_{\bm{0, \epsilon \beta}} & H_{\bm{0, \epsilon \epsilon}} \\
\end{bmatrix}
\begin{bmatrix}
\Delta \bm{\beta} \\
\Delta \bm{\epsilon}
\end{bmatrix}
\label{Taylor_expansion}
where the Jacobian vector $J_{\bm{0}}$ and the Hessian matrix $H_{\bm{0}}$ have been arranged in the partitioned form, and their subscript 0 denotes they are evaluated at $(\bm{\beta}_0, \bm{\epsilon}_0)$.
We then substitute this result into Eq. (\ref{dynamics_assumption}) to arrive at:
x_d(\bm{\beta}, \bm{\epsilon}) \approx
\Delta \bm{\beta}^T H_{\bm{0, \beta \epsilon}} \Delta \bm{\epsilon}
+ f_d(\bm{\beta}_0, \bm{\epsilon}_0) + c_d
\\ + 
\frac{1}{2} \Delta \bm{\beta}^T H_{\bm{0, \beta \beta}} \Delta \bm{\beta} 
+ J_{\bm{0, \beta}} \Delta \bm{\beta}
+ g_d(\bm{\beta}) 
+ \frac{1}{2} \Delta \bm{\epsilon}^T H_{\bm{0, \epsilon \epsilon}} \Delta \bm{\epsilon}
+ J_{\bm{0, \epsilon}} \Delta \bm{\epsilon}
+ h_d(\bm{\epsilon}) 
\label{approximation}
Comparing Eq. (\ref{approximation}) with Eq. (\ref{dynamics_assumption}), we can see our approximation factorizes the correlated term. If the same policy is executed in two different environment, the difference in the dynamics will be:
x_d(\bm{\beta}, \bm{\epsilon}_1) - 
x_d(\bm{\beta}, \bm{\epsilon}_2) \approx
\Delta \bm{\beta}^T H_{\bm{0, \beta \epsilon}} 
(\bm{\epsilon}_1 - \bm{\epsilon}_2)
+ h^*_d(\bm{\epsilon}_1) - h^*_d(\bm{\epsilon}_2) 
\text{where \,}
h^*_d(\bm{\epsilon}) = 
\frac{1}{2} \Delta \bm{\epsilon}^T H_{\bm{0, \epsilon \epsilon}} \Delta \bm{\epsilon}
+ J_{\bm{0, \epsilon}} \Delta \bm{\epsilon}
+ h_d(\bm{\epsilon}) 
\label{dynamics_diff}
If we treat the $h^*_d(\bm{\epsilon})$ term as an additional dimension of the environment vector, then Eq. (\ref{dynamics_diff}) can be represented as:
\Delta \bm{x}_d = 
\begin{bmatrix}
\Delta x_{d, 1} \\
\vdots \\
\Delta x_{d, N} \\
\end{bmatrix}
\approx
\begin{bmatrix}
\Delta \bm{\beta}_1^T H_{\bm{0, \beta \epsilon}} & 1 \\
\vdots & \vdots \\
\Delta \bm{\beta}_N^T H_{\bm{0, \beta \epsilon}} & 1 \\
\end{bmatrix}
\Delta \bm{\varepsilon} 
= W_{d}^T \Delta \bm{\varepsilon}
\label{linear_form}
where $\bm{\varepsilon}$ is our new environment vector. 
The vector $\Delta \bm{x}_d$ stands for the difference in dynamics for the entire repertoire with each entry $\Delta x_{d, i}$ denoting the result for the each elementary policy.
Combined with the assumption we made about the simulation, Eq. (\ref{linear_form}) suggests we may approximate the real-world distortion of the entire repertoire using a linear model.

The above derivations can be extended to any order of Taylor expansion, but the dimension of the new environment vector $\bm{\varepsilon}$ increases exponentially after the second order.
If we assume the dynamics can be efficiently approximated with the second order expansion, Eq. (\ref{linear_form}) gives a very potent result.
However, we do not have access to the matrix $W_d$ or the environment vector $\bm{\varepsilon}$ directly. 
Instead, we can leverage the fact that any environment vector can be linearly represented by a set of environment vectors as the basis.
Hence, if we have collected enough distortions, any new distortion can be  represented with a linear combination of the collected ones:
\begin{equation}
\begin{gathered}
\Delta \bm{x}_d^* \approx
W_{d}^T \Delta \bm{\varepsilon}^* = 
W_{d}^T
\begin{bmatrix}
\Delta \bm{\varepsilon}^{(1)}, \cdots, \Delta \bm{\varepsilon}^{(n)}
\end{bmatrix}
\begin{bmatrix}
\omega_1^* \\
\vdots \\
\omega_n^* \\
\end{bmatrix}
\\ =
\begin{bmatrix}
\Delta \bm{x}_d^{(1)}, \cdots, \Delta \bm{x}_d^{(n)}
\end{bmatrix}
\begin{bmatrix}
\omega_1^* \\
\vdots \\
\omega_n^* \\
\end{bmatrix}
\end{gathered}
\label{linear_combination}
\end{equation}
This means we need to get the distortions of all the elementary policies in several environments.
Since typically a repertoire can contain of a few thousand policies,
we cannot afford to do such amount of evaluations in the real-world, and hence we need to rely on simulation.

A question is that how many distortions we should collect.
If we get too few distortions, we might not have collected all the basis, and the linear model will perform poorly in uncovered environments.
On the other hand, if we have collected too many, not only will this be computationally expensive but this will also lead to very large dimensions. 
The solution is collect as many as we can afford and then reduce the dimensionality with singular vector decomposition (SVD). The collection of distortions can be represented as:
\begin{equation}
\bm{A} = 
\begin{bmatrix}
\Delta \bm{x}_d^{(1)}, \cdots, \Delta \bm{x}_d^{(n)}
\end{bmatrix}
= \sum_{i=1}^s \sigma_i \bm{v}_i \bm{u}_i^T
\label{SVD}
\end{equation}
where $\bm{v}_i$ stands for each normalized base vector and $\bm{u}_i$ is the normalized component of this base vector in each distortion.
The number of basis $s$ is much smaller than $n$ as we have collected more than we need to ensure a good coverage.
To conduct SVD, we left multiply matrix $\bm{A}$ by its transpose:
\begin{equation}
\begin{gathered}
\bm{A}^T \bm{A} = 
\sum_{i=1}^s \sigma_i^2 \bm{u}_i \bm{u}_i^T = \bm{U} \bm{D} \bm{U}^T
\\
\text{where \,}
\bm{U} = 
\begin{bmatrix}
\bm{u}_1, \cdots, \bm{u}_s
\end{bmatrix}
\text{, \,}
\bm{D} = 
\begin{bmatrix}
\sigma_1^2 &  &  \\
 & \ddots &  \\
 &  & \sigma_s^2 \\
\end{bmatrix}
\end{gathered}
\label{ATA}
\end{equation}
According to spectral theorem, entry $\sigma^2$ and vectors $\bm{u}$  are the non-zero eigenvalues and the corresponding eigenvectors of $\bm{A}^T\bm{A}$, respectively. 
The desired base vectors can then be calculated:
\begin{equation}
\begin{gathered}
%\frac{1}{\sigma_j}\bm{A} \bm{u}_j = 
%\sum_{i=1}^s \frac{\sigma_i}{\sigma_j} \bm{v}_i \delta_{i,j}
%= \bm{v}_j
%\\
%\text{hence \,}
\begin{bmatrix}
\bm{v}_1, \cdots, \bm{v}_s
\end{bmatrix} = 
\sqrt{\bm{D}^{-1}} \bm{A} \bm{U}
\end{gathered}
\label{basis}
\end{equation}
The eigenvalues denote the importance of each base vector.
If an eigenvalue is zero, we can completely ignore this dimension as it can be linearly represented by other basis.
In practice, since evaluations are noisy and our linear relation is approximated, we will never get any zero eigenvalues. 
To continue with our dimensionality reduction, we can rearrange the eigenvalues in a descending order and keep only the first few terms depending on how many dimensions we wish to have.
Thus, we can approximate any real-world distortion using a linear combination of simulated distortions filtered by SVD.
Although this is just an approximation, it should serve well as the prior mean function of GP.

\subsection{Determine the Input Space}
The GP requires the input space for the kernel to build the covariance matrix.
In the RTE, this is the same as the $x$ and $y$ displacements.
Considering the policies are generated using the MAP-Elites instead of gradient based methods, policies having similar displacements might have very different behaviours.
As a result, they may lead to very different outcomes in new environments, and the use of such input space could be misleading.
Hence, we aim to find the input space in which the distances between policies are related to their outcomes in all environments.

By examining our assumption of dynamics in Eq. (\ref{linear_form}), we see that the behaviour vectors that are encoded in the $W_d$ matrix.
By rearranging Eq. (\ref{linear_form}), we can get:
\begin{equation}
\begin{gathered}
W_{d}^T \Delta \bm{\varepsilon} = 
\begin{bmatrix}
\Delta \bm{\beta}_1^T & 1\\
\vdots & \vdots \\
\Delta \bm{\beta}_N^T & 1\\
\end{bmatrix}
%
\begin{bmatrix}
H_{\bm{0, \beta \epsilon}} & 0 \\
\\
0 & 1 \\
\end{bmatrix}
\Delta \bm{\varepsilon}
\end{gathered}
\label{linear_form_for_beta}
\end{equation}
Since we have found the basis that can linearly represent any environment vector $\bm{\varepsilon}$, we can always find a particular set of $\bm{\varepsilon}$ such that:
\begin{equation}
\begin{gathered}
W_{d}^T
\begin{bmatrix}
\Delta \bm{\varepsilon}_1, \cdots , \Delta \bm{\varepsilon}_N
\end{bmatrix}
= \\
\begin{bmatrix}
\Delta \bm{\beta}_1^T & 1\\
\vdots & \vdots \\
\Delta \bm{\beta}_N^T & 1\\
\end{bmatrix}
%
\begin{bmatrix}
H_{\bm{0, \beta \epsilon}} & 0 \\
\\
0 & 1 \\
\end{bmatrix}
\begin{bmatrix}
H_{\bm{0, \beta \epsilon}}^T
(H_{\bm{0, \beta \epsilon}}H_{\bm{0, \beta \epsilon}}^T)^{-1} \\
\\
0 \\
\end{bmatrix}
\\ 
= 
\begin{bmatrix}
\Delta \bm{\beta}_1^T & 1\\
\vdots & \vdots \\
\Delta \bm{\beta}_N^T & 1\\
\end{bmatrix}
%
\begin{bmatrix}
\mathbf{I} \\
\\
0 \\
\end{bmatrix} 
= 
\begin{bmatrix}
\Delta \bm{\beta}_1^T \\
\vdots \\
\Delta \bm{\beta}_N^T \\
\end{bmatrix}
\end{gathered}
\label{beta}
\end{equation}
This means we are able to represent the behaviour vectors as a linear combination of the distortion basis we found for the linear prior mean function.
This is a very reasonable result since if two policies have similar outcomes across many environments, they are more likely to share similar distortions in a new environment.
Note that the above derivations only hold if the dimension of the behaviour vector is not larger than the dimension of environment vector, otherwise the $H_{\bm{0, \beta \epsilon}}H_{\bm{0, \beta \epsilon}}^T$ matrix is not invertible.
However, our purpose of finding the behaviour vectors is just to build statistical inference between policies, which tolerates uncertainties.
In the case there are omitted dimensions, we can just add a larger noise to the diagonal of the covariance matrix to incorporate this missing information.

It is hard to determine the detailed linear combination to get the behaviour vectors.
But considering the use of anisotropic length scales for the distance metric in Eq. (\ref{distance_metric}), it is desirable to find a linear combination to reduce the correlations between the dimensions.
This is achieved using principle component analysis (PCA) \cite{PCA}. We first find the empirical covariance matrix of the standardized distortions.
\begin{equation}
\begin{gathered}
\bm{\Sigma} = \bm{X}^T \bm{X}
\end{gathered}
\label{cov}
\end{equation}
$\bm{X}$ is a $N \times s$ matrix, where $N$ is the size of the repertoire (the amount of elementary policies) and $s$ is the number of basis found using SVD.
Each column of $\bm{X}$, namely each dimension of basis, is standardized across the policies.
The covariance matrix $\bm{\Sigma}$ is a real-symmetric matrix, according to spectral theorem, it can be represented by:
\begin{equation}
\begin{gathered}
\bm{\Sigma} = 
\sum_{i=1}^s \lambda_i \bm{u}_i \bm{u}_i^T = \bm{U} \bm{D} \bm{U}^T
\end{gathered}
\label{PCA}
\end{equation}
Similar to SVD, the $\lambda_i$ are the eigenvalues of the covariance matrix, and $\bm{U}$ contains the corresponding eigenvectors.
%Similar to SVD, these eigenvalues are non-negative and give the importance of each dimension (see Fig. \ref{}).
We can then find our input space with the following transform:
\begin{equation}
\begin{gathered}
\begin{bmatrix}
\bm{v}'_1, \cdots, \bm{v}'_s
\end{bmatrix}
=
\begin{bmatrix}
\bm{v}_1, \cdots, \bm{v}_s
\end{bmatrix}
\bm{U} \sqrt{\bm{D}^{-1}}
\end{gathered}
\label{input_space}
\end{equation}
This gives the coordinates of each elementary policy in our input space. Each dimension of the coordinates are now decorrelated (see Fig. \ref{PCA_outcome}).
%
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.255\textwidth]{before_pca.pdf}%
\label{fig_first_case}}
%\hfil
\subfloat[]{\includegraphics[width=0.195\textwidth]{after_pca.pdf}%
\label{fig_second_case}}
\caption{Example of PCA transform. Original data distribution is given by (a), where the two axes are highly correlated. (b) is the resultant distribution after applying Eq. (\ref{input_space}), where the correlation is removed.}
\label{PCA_outcome}
\end{figure}

>>>>

\usepackage{bm}
%\usepackage{amssymb}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\renewcommand{\algorithmiccomment}[1]{\hfill $\triangleright$ #1}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\hypersetup{hidelinks}
%\usepackage[colorlinks, linkcolor=blue, citecolor=green, urlcolor=red]{hyperref}

@@

%\begin{multicols}{2}
\input{introduction.tex}

@@

\begin{algorithm}
\caption{MAP-Elites}
\begin{algorithmic}
\STATE \textbf{procedure} MAP-Elites
\STATE Discretize the task space into grids
\STATE $\mathcal{R} \leftarrow \emptyset$ \COMMENT{Initialize repertoire}
\FOR{$\bm{i} = 1 \rightarrow G$}
\STATE $\theta \leftarrow $ generate{\_}random{\_}policy()
\STATE $\Theta$, $f$ $\leftarrow$ evaluate{\_}policy($\theta$) \COMMENT{Get outcome and fitness}
\STATE add{\_}to{\_}repertoire($\Theta$, $f$, $\theta$, $\mathcal{R}$) \COMMENT{Update repertoire}
\ENDFOR
\FOR{$\bm{i} = 1 \rightarrow T$}
\STATE $\vartheta \leftarrow$ select{\_}parents($\mathcal{R}$)
\STATE $\bm{\theta} \leftarrow$ crossover($\vartheta$, $N$) \COMMENT{Generate $N$ off-springs}
\FOR{$\bm{j} = 1 \rightarrow N$}
\STATE $\Theta$, $f$ $\leftarrow$ evaluate{\_}policy($\bm{\theta_j}$)
\COMMENT{Off-spring evaluation}
\STATE add{\_}to{\_}repertoire($\Theta$, $f$, $\bm{\theta_j}$, $\mathcal{R}$)
\ENDFOR
\ENDFOR
\end{algorithmic}
\label{MAP-Elites}
\end{algorithm}
\noindent
Evaluating the policies in the simulation is the major source of computationally expense for the MAP-Elites.
A merit of this algorithm is that the inner for loop of off-spring evaluation can be conducted in parallel, and the repertoire update can be made after having collected all the results of the off-springs.
This enables us to generate a large number of off-springs in each iteration and evaluate them in parallel, fully leveraging the power of modern Cluster computing \cite{Q-Dax}.
By running the algorithm for a few thousand iterations, we will end up having a repertoire of high-performing policies well-covering the task space.

\subsection{Gaussian Process}
The Gaussian Process (GP) is a powerful non-parametric machine learning algorithm. 
It assumes the prior distribution of the outcomes to be a multivariate normal distribution with a prior mean $\mu$ and a covariance matrix $\Sigma$. 
\begin{equation}
f(\bm{y}) = \frac{exp(
-\frac{1}{2}(\bm{y} - \bm{\mu})^T \bm{\Sigma^{-1}} (\bm{y} - \bm{\mu})
)}
{
\sqrt{(2\pi)^N \det(\bm{\Sigma})}
}
\label{multi_normal}
\end{equation}
In contrast to regular multivariate normal distribution, GP allows the prior mean and the covariance matrix to be input dependent. The
prior mean becomes a function $\mu(\bm{x})$, and the covariance between two data points is given by a covariance function 
$k(\bm{x_1}, \bm{x_2})$ called kernel. 
Following the fact that data points closer to each other are more correlated, this covariance starts from the prior variance and asymptotically decreases to zero as the distance between the two points tends to infinity. 
To account for the fact the observations might be noisy, we can incorporate the noise into the covariance matrix by adding the variance of the noise on the diagonal.
\begin{equation}
\bm{\Sigma} = \bm{K}(\bm{x_{1:N}}, \bm{x_{1:N}}) + \sigma_n^2 \mathbf{I}
\label{covariance_matrix}
\end{equation}

This allows us to build reasonable joint prior distribution for the values of any well-behaved functions.
To make prediction with GP, we first use the prior mean function and kernel to build up the joint distribution as the prior. 
Then we can calculate the posterior for the data we wish to evaluate conditioned on the ones we have observed using Bayes rule.
Since the Gaussian distribution is a conjugate distribution, the posterior is still a Gaussian with mean and variance given by \cite{GP, GP_posterior}:
\begin{equation}
\begin{gathered}
\mu_*(x) = \mu(x) + 
\bm{\Sigma}_{N, x}^T
\bm{\Sigma}_{N, N}^{-1}
(\bm{y}_{1:N} - \bm{\mu}(\bm{x}_{1:N}))
\\
\sigma^2_*(x) = \sigma^2(x) -  
\bm{\Sigma}_{N, x}^T
\bm{\Sigma}_{N, N}^{-1}
\bm{\Sigma}_{N, x}
\end{gathered}
\label{GP_posterior}
\end{equation}
where $\bm{\Sigma}_{N, N}$ denotes the covariance matrix of the observed data, and $\bm{\Sigma}_{N, x}$ denotes the column vector with entry on each row equal to the covariance between 
each observed data and the data we hope to evaluate.

@@

\input{methodology.tex}

%To formulate the problem clearly, we are aiming to leverage the collaboratively collected real-world interactions to assist a robot deployed in an unknown environment to quickly adapt and complete its task.
%We are provided with some episodes (each task-solving process corresponds to an episode of interaction data) of real-world interactions collected by several robots deployed in an unknown distribution of environments.
%The most intuitive idea is to label these historical data with their sources (like collected from grass terrain, collected with left leg broken, etc.) and build a simple database in the cloud server. 
%During the adaptation, the robot examines the current environment and uploads the result to the database server. The server then identifies the most similar environment recorded in the database and let the robot download the corresponding data to assist the its adaptation.
%However, accurately measuring and describing the environment can be expensive and difficult. It is also hard to determine whether a previously encountered environment shares similar dynamics with the current one even if we have accurate descriptions for them.
%In the case we do know that some historical data come from exactly the same environment as the current one, these data might be insufficient in quantity to assist the current task.
%Hence, it is reasonable to group the episodes resulting from similar dynamics into clusters, ensuring having enough data for analysis.
%We will assume that we don’t have access to any information of the environment and aim to conduct clustering only according to the dynamics.

\input{experiment.tex}
